---
title: "2.5 Functions"
---

Open the slides in a new tab [here](slides/05-functions.qmd){target="_blank"}.

<div>

```{=html}
<iframe class="slide-deck" src="slides/05-functions.html"></iframe>
```

</div>

## Exercises

1. Square a number. You're tired of writing `x^2` when you want to square `x`, so you want a function to square a number. You can call it `square()`. I showed this in the slides, now try on your own!

```{r, eval = FALSE}
# start out with a number to test
x <- 3
# you'll want your function to return this number
x^2
square <- function() {
  
}
# test it out
square(x)
square(53)
53^2 # does this match?
```

2. Raise to any power. You don't just want to square numbers, you want to raise them to higher powers too. Make a function that uses two arguments, `x` for a number, and `power` for the power. Call it `raise()`.

```{r, eval = FALSE}
raise <- function() {
  
}

# test with
raise(x = 2, power = 4)
# should give you
2^4
```

3. Change your `raise()` function to default to squaring `x` when the user doesn't enter a value for `power`.

```{r, eval = FALSE}
# test
raise(x = 5)
# should give you
5^2
```

## Resources

-   [Advanced R](-	https://adv-r.hadley.nz/functions.html)